% Core Data Entities Section
% Placeholder - will be enhanced with entity tables

\subsection{Core Traceability Entities}
\label{sec:core-traceability-entities}

Implementations claiming BOOST Core conformance \MUST{} support these entities:

\subsubsection{TraceableUnit}
\label{sec:entity-traceable-unit}

The central entity in BOOST's hub-and-spoke architecture. Every TRU connects to \secref{sec:entity-organization}, \secref{sec:entity-material}, and \secref{sec:entity-geographic-data} entities.

\begin{normative}[title=TraceableUnit Requirements]
Every \TRU{} \MUST{} have:
\begin{itemize}
    \item Unique identifier with biometric signature capability
    \item Material type classification
    \item Volume and mass measurements
    \item Geographic location data
    \item Processing history linkage
\end{itemize}
\end{normative}

% Entity table will be generated from schema files
\input{tex/entities/traceable-unit-table}

\subsubsection{Organization}
\label{sec:entity-organization}

Defines harvesters, processors, and supply chain participants. Referenced by \TRU{} entities and connects to Certificate entities for certification management.

% Entity table will be generated from schema files
\input{tex/entities/organization-table}

\subsubsection{Material}
\label{sec:entity-material}

Material type classifications referenced by \TRU{} entities. For multi-species materials, connects to SpeciesComponent entities.

% Entity table will be generated from schema files
\input{tex/entities/material-table}

\subsubsection{GeographicData}
\label{sec:entity-geographic-data}

Spatial data in GeoJSON format for location tracking. Referenced by \TRU{} and LocationHistory entities for comprehensive location management.

% Entity table will be generated from schema files
\input{tex/entities/geographic-data-table}